August 13th, 2022

Grouping and summarizing
Summarizing the median life expectancy
```{r}
gapminder %>%
  summarize(medianLifeExp = median(lifeExp))

#Filter for 1957 then summarize the median lifeExp
gapminder %>%
  filter(year==1957) %>%
  summarize(medianLifeExp = median(lifeExp))

gapminder %>%
  filter(year==1957) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap=max(gdpPercap))

```

Summarizing by year

```{r}
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```
Summarizing by continent
```{r}
gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap)) %>%
  arrange(desc(maxGdpPercap))
```
Summarizing by continent and year
```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```
Visualizing median life expectancy over time
```{r}
by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_year

ggplot(by_year, aes(x=year, y=medianGdpPercap)) +
  geom_line()
```
```{r}
by_year_continet <- gapminder %>%
  group_by(continent, year) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_year_continet

ggplot(by_year_continet, aes(x=year, y=medianGdpPercap, color=continent)) + 
  geom_line()
```
Bar Plots
Visualizing median GDP per capita by continent
```{r}
by_continent <- gapminder%>%
  filter(year==1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

by_continent

ggplot(by_continent, aes(x=continent, y=medianGdpPercap)) +
  geom_col()
```
Visualizing GDP per capita by country in Oceania
```{r}
oceania_1952 <- gapminder %>%
  filter(continent == "Oceania", year == 1952)

oceania_1952

ggplot(oceania_1952, aes(x=country, y=gdpPercap)) + 
  geom_col()
```
Histogram
```{r}
library(gapminder)
library(ggplot2)
library(dplyr)

gapminder

gapminder_1952 <- gapminder %>%
  filter(year==1952) %>%
  mutate(pop_by_mil=pop/1000000)

ggplot(gapminder_1952, aes(x=pop_by_mil)) +
  geom_histogram(bins = 50)
```
visualizing population with x-axis on a log scale

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop))+
  geom_histogram(bins=50) +
  scale_x_log10()
```
Boxplots
Comparin GDP per capita across continents
```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() +
  scale_y_log10()
```
Adding a title to your graph
```{r}
ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() + 
  ggtitle("Comparing GDP per capita across continents")
```
Data manipulation with dplyr
```{r}
library(tidyverse)

counties <- read_csv('counties.csv')
glimpse(counties)

counties
```
Selecting columns
```{r}

counties %>%
  select(state, county, population, men, women)

counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected %>%
  arrange(desc(public_work))

counties_selected <- counties %>%
  select(state, county, population)

counties_selected %>%
  filter(population > 1000000)

counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected %>%
  filter(state=="Texas", population>10000) %>%
  arrange(desc(private_work))
```
Select, mutate, filter and arrange
```{r}
counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(population > 10000) %>%
  arrange(desc(proportion_men))
```
The verb count
```{r}
select_counties <- counties %>%
  select(state, region, county, population, citizens)

counties %>%
  count(region, sort=TRUE) %>%
  arrange(desc(n))
```
Counting citizens by state
```{r}
counties_selected <- counties %>%
  select(county, region, state, population, citizens)

counties_selected %>%
  count(state, wt=citizens) %>%
  arrange(desc(n))
```
Mutating and counting
```{r}
counties_selected <- counties %>%
  select(county, region, state, population, walk)

counties_selected %>%
  mutate(population_walk = population*walk / 100) %>%
  count(state, wt=population_walk, sort=TRUE)
```
Summarizing
```{r}
counties_selected <- counties %>%
  select(county, population, income, unemployment)


counties_selected %>%
  summarize(min_population=min(population), max_unemployment=max(unemployment), averange_income=mean(income))
```
Summarizing by state
```{r}
counties_selected <- counties %>%
  select(state, county, population, land_area)

counties_selected %>%
  group_by(state) %>%
  summarize(total_area=sum(land_area), total_population = sum(population)) %>%
  mutate(density=total_population/total_area) %>%
  arrange(desc(density))

counties_selected <- counties %>%
  select(region, state, county, population)

counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop = sum(population)) %>%
  summarize(averange_pop=mean(total_pop), median_pop=median(total_pop))
```
Top_n verb
```{r}
counties_selected <- counties %>%
  select(region, state, county, metro, population, walk)

counties_selected %>%
  group_by(region) %>%
  top_n(1, walk)

counties_selected <- counties %>%
  select(region, state, county, population, income)

counties_selected %>%
  group_by(region, state) %>%
  summarize(averange_income=mean(income)) %>%
  top_n(1,averange_income)
```
```{r}
counties_selected <- counties %>%
  select(state, metro, population)

counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop=sum(population)) %>%
  top_n(1, total_pop) %>%
  ungroup() %>%
  count(metro)
```



